---
title: "Alternate Transformation - New Public/Private Feature"
author: "Ryan Rogers"
date: '2022-10-10'
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(caret)
library(MLmetrics)
library(rpart)
```

# Concept
The idea behind this notebook is to determine if the process of converting the initial input format (with columns for public and private data) into a new format (with columns for mean/median values and a public/private indicator) yields better results with simple regression models. No other supplementary features will be included at this time. Standard filtering of rows (remove private rows with invalid counts, etc. will be applied.)

## Data Load and Transformation

### Data Load

```{r}
original_data <- read.csv("priv_mcare_f_pay.csv")
```


### Data Filtering

Note: I presume that all data from forbidden MSAs is off limits. Therefore, we will simply drop all rows where priv_count is NA or 0

```{r}
filtered_data <- original_data %>%
  filter(!is.na(priv_count) & (priv_count > 0) & !is.na(lon) & !is.na(lat))
```

### Data Preprocessing

One thing we will do before much of the transformation is deal with the categorical variables.

```{r}
filtered_data_important_fields <- filtered_data %>% select(!c(CBSA_NAME, FIPS.State.Code))
cat_encoder <- dummyVars(" ~ .", data=filtered_data_important_fields)
cat_encoded <- data.frame(predict(cat_encoder, filtered_data_important_fields))

cat_encoded$index <- 1:nrow(cat_encoded)
```

### Data Splitting

For filters applied, we have:

* Public (mcare):
  + Drop all NAs
* Private:
  + Drop all NAs
  + Drop all rows with no mcare_los
  + Drop all rows with priv_count < 50
  
For general transformations:

* public_private column added
* SD/IQR columns dropped. They aren't exactly comparable
* Columns renamed for dataset recombination.

```{r}
filtered_data_public <-
  cat_encoded %>%
  select(!c(priv_count, priv_pay_iqr, priv_pay_mean, priv_pay_median, mcare_pay_sd)) %>%
  mutate(public_private = 'public') %>%
  rename(pay_mean = mcare_pay_mean) %>%
  rename(pay_median = mcare_pay_median) %>%
  filter(!is.na(pay_mean))

filtered_data_private <-
  cat_encoded %>%
  filter(priv_count >= 50) %>%
  select(!c(priv_count, priv_pay_iqr, mcare_pay_mean, mcare_pay_median, mcare_pay_sd)) %>%
  mutate(public_private = 'private') %>%
  rename(pay_mean = priv_pay_mean) %>%
  rename(pay_median = priv_pay_median) %>%
  filter(!is.na(pay_mean) & !is.na(mcare_los))
```

### Separate out test set

I arbitrarily grabbed 25% of the private records post-filtering (we are only interested in predicting using these).

```{r}
test_set <-
  filtered_data_private %>%
  sample_frac(0.25)

filtered_data_private <- anti_join(filtered_data_private,
                                   test_set,
                                   by = ('index'))
```


### Data Recombination

```{r}
dev_set <- rbind(filtered_data_public, filtered_data_private)
```

### Create Development and Test Sets for Original Data

```{r}
untransformed_data <- cat_encoded %>%
  filter((priv_count >= 50) & !is.na(mcare_los)) %>%
  select(!c(priv_pay_iqr, mcare_pay_sd))

untransformed_test_set <-
  untransformed_data %>%
  sample_frac(0.25)

untransformed_dev_set <- anti_join(untransformed_data,
                                   untransformed_test_set,
                                   by = ('index'))
```


## Comparing Performance

### Original Dataset

#### Linear Regression

```{r}
orig_lm <- lm(formula = priv_pay_median ~ ., data = (untransformed_dev_set %>% select(!c(priv_pay_mean, msa, index))))
#train(
#  priv_pay_median ~ .,
#  data = (untransformed_dev_set %>% select(!c(priv_pay_mean, msa, index))),
#  method = 'lasso'
#)
summary(orig_lm)

orig_lm_pred <- predict(orig_lm, newdata = untransformed_test_set)

print("")
print("MAPE is:")
MAPE(orig_lm_pred, untransformed_test_set$priv_pay_median)
```
#### Decision Tree Regression

```{r}
orig_tree <- rpart(formula = priv_pay_median ~ ., data = (untransformed_dev_set %>% select(!c(priv_pay_mean, msa, index))))
summary(orig_tree)

orig_tree_pred <- predict(orig_tree, newdata = untransformed_test_set)

print("")
print("MAPE is:")
MAPE(orig_tree_pred, untransformed_test_set$priv_pay_median)
```


### Transformed Dataset

#### Linear Regression

```{r}
transformed_lm <- lm(formula = pay_median ~ ., data = (dev_set %>% select(!c(pay_mean, msa, index))))
#train(
#  priv_pay_median ~ .,
#  data = (untransformed_dev_set %>% select(!c(priv_pay_mean, msa, index))),
#  method = 'lasso'
#)
summary(transformed_lm)

transformed_lm_pred <- predict(transformed_lm, newdata = test_set)

print("")
print("MAPE is:")
MAPE(transformed_lm_pred, test_set$pay_median)
```

#### Decision Tree Regression

```{r}
transformed_tree <- rpart(formula = pay_median ~ ., data = (dev_set %>% select(!c(pay_mean, msa, index))))
summary(transformed_tree)

transformed_tree_pred <- predict(transformed_tree, newdata = test_set)

print("")
print("MAPE is:")
MAPE(transformed_tree_pred, test_set$pay_median)
```

