---
title: "Additional Dataset Exploration"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
business_data <- read.csv("business_msa_data.csv")
data <- read.csv("priv_pay_data_with_hospital.csv")
econ_data <- read.csv("ACSST5Y2020.S1811-2022-10-18T150043.csv")

econ_data_long <- econ_data %>%
  filter(Label..Grouping. == "Population Age 16 and Over") %>%
  pivot_longer(cols = c(2:2818),
               names_to = "MSA",
               values_to = "Population")

split_data <- econ_data_long %>%
  filter(!grepl("Micro",MSA)) %>%
  separate(col = MSA, sep = "Metro", into = c("MSAName", NA)) %>%
  mutate(Population = as.numeric(gsub(",", "", Population))) %>%
  group_by(MSAName) %>%
  summarise(Population = max(Population))

msas_in_data <- data %>%
  select(msa, CBSA_NAME) %>%
  distinct()

msa_list <- read.csv("msa_list.csv",header = FALSE)
msa_list2 <- msa_list %>%
  slice(4:1923) %>%
  filter(V5 == "Metropolitan Statistical Area") %>%
  select(V1,V4) %>%
  distinct() %>%
  rename(msa = V1,
         MSA_Name = V4) %>%
  select(-msa)

msa_list3 <- msa_list %>%
  slice(4:1923) %>%
  filter(V5 == "Metropolitan Statistical Area") %>%
  select(V1,V2,V4) %>%
  mutate(V1 = ifelse(V2 != "", V2, V1)) %>%
  distinct() %>%
  rename(msa = V1,
         MSA_Name = V4) %>%
  select(-V2) %>%
  mutate(msa = as.numeric(msa))

pop_data <- cbind(msa_list2, split_data)

pop_data2 <- left_join(msa_list3, pop_data, by = "MSA_Name")

  

```


# Population Data (S1811)


```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA}
msa_pay_data <- data %>%
  group_by(msa,CBSA_NAME) %>%
  summarise(msa_priv_pay_mean = weighted.mean(priv_pay_mean, priv_count, na.rm = T),
            msa_priv_pay_median = median(priv_pay_median, priv_count, na.rm = T))

full_data_with_pop <- left_join(msa_pay_data, pop_data2, by = "msa")

full_data_with_pop2 <- full_data_with_pop %>%
  select(msa_priv_pay_mean, msa_priv_pay_median, Population)

cor(full_data_with_pop2, use = "complete.obs")

ggplot(data = full_data_with_pop2, aes(x = Population, y = msa_priv_pay_mean)) + geom_point() +
  labs(title = "Private Payout Mean by Population of MSA",
       x = "Total Population",
       y = "Private Payout Mean (by MSA)")


```

* Not seeing a ton of correlation between MSA population and private payouts
* Might be worth including in the model as it could still be significant
* Likely a better variable to include rather than employment or payroll

# Population Data (S1811) by Group


```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA}
msa_pay_data <- data %>%
  group_by(msa,CBSA_NAME,group) %>%
  summarise(msa_group_priv_pay_mean = weighted.mean(priv_pay_mean, priv_count, na.rm = T),
            msa_group_priv_pay_median = median(priv_pay_median, priv_count, na.rm = T))

full_data_with_pop <- left_join(msa_pay_data, pop_data2, by = "msa")

full_data_with_pop2 <- full_data_with_pop %>%
  ungroup() %>%
  select(msa_group_priv_pay_mean, msa_group_priv_pay_median, Population)

cor(full_data_with_pop2, use = "complete.obs")


```

* Not seeing a strong correlation between population and private payment at MSA/group level


# Total Employment and Annual Payroll

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA}

total_employment <- business_data %>%
  filter(naics == "------") %>%
  select(msa, emp, ap)

msa_pay_data <- data %>%
  group_by(msa) %>%
  summarise(msa_priv_pay_mean = weighted.mean(priv_pay_mean, priv_count, na.rm = T),
            msa_priv_pay_median = median(priv_pay_median, priv_count, na.rm = T))

full_data <- left_join(msa_pay_data, total_employment, by = "msa")

cor(full_data, use = "complete.obs")

ggplot(data = full_data, aes(x = emp, y = msa_priv_pay_mean)) + geom_point() +
  labs(title = "Private Payout Mean by Total Employment of MSA",
       x = "Total Employment",
       y = "Private Payout Mean (by MSA)")
ggplot(data = full_data, aes(x = ap, y = msa_priv_pay_mean)) + geom_point() +
  labs(title = "Private Payout Mean by Annual Payroll of MSA",
       x = "Annual Payroll",
       y = "Private Payout Mean (by MSA)")




```


* Not seeing much correlation between MSA employment or annual payroll with private payments
* Seems more logical to use population rather than employment or annual payroll if we choose to use this type of feature in the model


# Total Employment and Annual Payroll at Group Level

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA}

total_employment <- business_data %>%
  filter(naics == "------") %>%
  select(msa, emp, ap)

msa_pay_data <- data %>%
  group_by(msa, group) %>%
  summarise(msa_priv_pay_mean = weighted.mean(priv_pay_mean, priv_count, na.rm = T),
            msa_priv_pay_median = median(priv_pay_median, priv_count, na.rm = T))

full_data <- left_join(msa_pay_data, total_employment, by = "msa")


full_data_group <- full_data %>%
  ungroup() %>%
  group_by(group) %>%
  summarise(group_priv_pay_mean = mean(msa_priv_pay_mean, na.rm = T),
            group_priv_pay_median = median(msa_priv_pay_median, na.rm = T),
            emp = mean(emp, na.rm = T),
            ap = mean(ap, na.rm = T))
  




```


* Correlation of payment by group varying depending on employment/payroll of MSA doesn't make sense to include