---
title: "Compartmentalized Baseline Modeling"
author: "Ryan Rogers"
date: "10/28/2022"
output: pdf_document
---

Library imports are left as-is. They'll be necessary in almost every version. New imports added for helper libraries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(randomForest)
library(MLmetrics)
source(file = "../Common_Functions/processing_functions.R") # import of data_split transformation
source(file = "../Common_Functions/baseline_model.R") # import of baseline model
```

## Modeling with threshold 50 number of claims

Will leave data import alone for now.

```{r, warning = FALSE, message = FALSE}
data <- read.csv("priv_mcare_f_pay_2022Oct18.csv")
hospital_data <- read.csv("Hospital_Master_Sheet.csv")
```

Hospital data aggregation, data split, and data filtering are now compartmentalized

```{r, warning = FALSE, message = FALSE}
# Hospital data aggregation - validated for sameness
hospitals_msa <- hospital_data %>% aggregate_hospital_features()
rm(hospital_data)

# Data split into model data and predict - varies from original slightly
split_dataset <- data %>% data_split(count_thresh = 49)
working_set <- split_dataset[[1]]
predict_set <- split_dataset[[2]]

model_data <- left_join(working_set, hospitals_msa, by = "msa")  %>%
  select(-priv_pay_mean, -priv_pay_iqr, -mcare_pay_mean, -mcare_pay_sd, -Urban, -msa)
rm(working_set)

predict_data <- left_join(predict_set, hospitals_msa, by = "msa")  %>%
  select(-priv_pay_mean, -priv_pay_iqr, -mcare_pay_mean, -mcare_pay_sd, -Urban, -msa)
rm(predict_set)

# Train test split
train_test_data <- model_data %>% train_test_split(proportion = 0.8)
rm(model_data)

train <- train_test_data[[1]]
test <- train_test_data[[2]]

```

Model Creation and Prediction are now compartmentalized

```{r, warning = FALSE, message = FALSE}
# Random Forest model

# Fit Random Forest Model on training data
Random_Forest <- baseline_rdm_forest(data = train)

train_predict <- make_baseline_prediction(Random_Forest, train)
rm(train)

train_mape_percent = get_mape_percentage(train_predict)
```


```{r, warning = FALSE, message = FALSE}
varImpPlot(Random_Forest, bg = "aquamarine3")

```


```{r, warning = FALSE, message = FALSE}
test_predict <- make_baseline_prediction(Random_Forest, test)
rm(test)

test_mape_percent = get_mape_percentage(test_predict)
```

```{r}
cat("With Threshold >50 claims for training set:\n")
cat("Train MAPE:" , round(train_mape_percent, 2), "%\n")
cat("Test MAPE:" , round(test_mape_percent, 2), "%\n")
```


## Modeling with threshold 35 number of claims

Will leave data import alone for now.

```{r, warning = FALSE, message = FALSE}
data <- read.csv("priv_mcare_f_pay_2022Oct18.csv")
hospital_data <- read.csv("Hospital_Master_Sheet.csv")
```

Hospital data aggregation, data split, and data filtering are now compartmentalized

```{r, warning = FALSE, message = FALSE}
hospitals_msa <- hospital_data %>% aggregate_hospital_features()
rm(hospital_data)

new_data <- data %>%
  filter(priv_pay_median >= 0 | is.na(priv_pay_median)) %>%
  filter(priv_count != 0)

new_data_with_hospital <- left_join(new_data, hospitals_msa, by = "msa")


model_data <- new_data_with_hospital %>%
  filter(priv_count >= 35) %>%
  filter(!is.na(priv_pay_median)) %>%
  select(-priv_pay_mean, -priv_pay_iqr, -mcare_pay_mean, -mcare_pay_sd,-Urban, -msa)

predict_data <- new_data_with_hospital %>%
  filter(priv_count <= 10) %>%
  filter(priv_pay_median > 0)

set.seed(123) #Set seed for reproducible analysis
dt = sort(sample(nrow(model_data), nrow(model_data)*.8)) #Split data
train <-model_data[dt,] #80% training data
test <-model_data[-dt,] #20% test data



# Random Forest model
set.seed(123) #Set seed for reproducibility 
# Fit Random Forest Model on training data
Random_Forest <- randomForest(
  formula = priv_pay_median ~ .,
  data    = train,
  num.trees = 500,
  mtry = 7,
  nodesize = 20,
  na.action = na.omit
)

train_predict <- train %>%
  mutate(pred_priv_pay_median = predict(Random_Forest, train)) %>%
  filter(!is.na(pred_priv_pay_median))
trn_m = MAPE(train_predict$pred_priv_pay_median, train_predict$priv_pay_median)
train_mape_percent = mean(abs((train_predict$priv_pay_median - train_predict$pred_priv_pay_median)/train_predict$priv_pay_median),na.rm = T)*100
```

```{r, warning = FALSE, message = FALSE}
varImpPlot(Random_Forest, bg = "aquamarine3")

```

```{r, warning = FALSE, message = FALSE}
test_predict <- test %>%
  mutate(pred_priv_pay_median = predict(Random_Forest, test)) %>%
  filter(!is.na(pred_priv_pay_median))
tst_m = MAPE(test_predict$pred_priv_pay_median, test_predict$priv_pay_median)
test_mape_percent = mean(abs((test_predict$priv_pay_median - test_predict$pred_priv_pay_median)/test_predict$priv_pay_median),na.rm = T)*100
```

```{r}
cat("With Threshold >35 claims for training set:\n")
cat("Train MAPE:" , round(train_mape_percent, 2), "%\n")
cat("Test MAPE:" , round(test_mape_percent, 2), "%\n")
```
